---
title: "BootLoader"
subtitle: "Who Wants a Bike?"
author: "Miles D. Williams"
date: "3/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=T,
  message=F,
  warning=F
)
options(digits=3)
```


First I open the `tidyverse` because that's how I roll...

```{r}
library(tidyverse)
```

Now I get the data and given it a nice short name (`dt`)

```{r}
dt = read_csv('dataviz-01.csv')
dt
```

How many unique people are in the data?

```{r}
dt %>%
  count(userid) %>%
  arrange(-n) %>%
  mutate(id=1:n()) %>%
  ggplot() +
  aes(id,n) +
  geom_col() +
  scale_y_continuous(breaks=0:25)
```

So it looks like a bunch of people are regulars. That'll be important to note for later.

We want to see which people are most likely to contribute money toward Dennis' 3-speed bike concept. What categories of projects to people contribute funds toward?

```{r}
dt %>%
  count(category)
```

It looks like we have 5 categories: environment, fashion, games, sports, and tech. 

Now, it might also be useful to look at the people that either only viewed a project vs. those who funded.

```{r}
dt %>%
  count(category, eventType)
```

A slight minority of people actually fund projects. 

When they do donate, how much do they contribute? Below I look at the distribution of payments by category.

```{r}
dt %>%
  filter(eventType == 'fundedProject') %>%
  ggplot() +
  aes(amountDonated) +
  geom_histogram() +
  facet_wrap(~category)
```


So, I think Dennis would obviously benefit most if the people who receive a push are those who are 

1. most likely to fund projects;

2. likely to given more than less money when they do contribute;

3. likely to contribute more than once;

4. likely to contribute to categories relevant to Dennis' project.

First, let's identify those who aren't so tight fisted. How many people don't contribute anything?

```{r}
dt %>%
  group_by(userid) %>%
  count(is.na(amountDonated)) %>%
  filter(n==0)
```

So everybody contributes something.

Let's just do this as a predictive model:

```{r}
library(randomForest)
```

Make NAs 0s

```{r}
dt$amountDonated[is.na(dt$amountDonated)]=0
dt = dt %>%
  mutate_if(
    is.character,
    as.factor
  ) %>%
  mutate(
    zipCode = as.factor(zipCode)
  )
# and let's log-transform money

dt = dt %>%
  mutate(lamount = log(amountDonated+1),
         funded = as.factor(amountDonated>0))
```


```{r}
startTime = Sys.time()
rf = randomForest(
  amountDonated ~
    eventType + category + 
    gender + age + maritalStatus + 
    device + state + city, 
  dt
)
endTime = Sys.time()
startTime - endTime # time to run: about 10 minutes.
```

```{r}
trainObs = sample(
  1:nrow(dt), replace = F, size = round(nrow(dt)*.7)
)
train_dt = dt[trainObs,] # 70% of obs.
test_dt = dt[-trainObs,] # 30% of obs.
```

```{r}
startTime = Sys.time()
rf1 = randomForest(
  funded ~
    eventType + category + 
    gender + age + maritalStatus + 
    device + state, 
  train_dt
)
endTime = Sys.time()
endTime - startTime # time to train
```

```{r}
preds = predict(rf1, newdata = test_dt)
ggplot() +
  aes(test_dt$funded,preds) +
  geom_point() +
  geom_smooth() +
  geom_abline(intercept = 0, slope = 1) +
  labs(
    caption = paste0(
      round(100*mean(preds==test_dt$funded),2),
      '% of variation predicted'
    )
  )
```

So now we have a REALLY GOOD predictive model. Now we just need to identify the people who are our likely funders...